Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-11-13 index.qmd
add Laurent's comments

plot_RAW <- RAW %>%  
  ggplot(aes(x=Low, y=High))+
  # geom_point(aes(color = as.factor(pred)))
  ggpointdensity::geom_pointdensity(adjust = .03)+
  viridis::scale_color_viridis(limits = c(0,6000),
                               breaks = c(0,2000,4000,6000),
                               labels = c("1","2000","4000","6000"))+
  # geom_smooth(method = "loess")+
  geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed", color = "black" )+
  # geom_smooth(method = "lm", color = "red", linetype = "dashed",linewidth = 1)+
  theme_Bede()+
  ylim(c(0.05,0.6))+
  xlim(c(0.05,0.6))+
  ylab("Reflectance of the 120m flight")+
  xlab("Reflectance of the 12m flight")+
  labs(color = "Pixel density")+
  theme(
    legend.position = "top",
        legend.key.width = unit(3,"line"))

plot_STD <- STD %>%  
    ggplot(aes(x=Low, y=High))+
  # geom_point(aes(color = as.factor(pred)))
  ggpointdensity::geom_pointdensity(adjust = .03)+
  viridis::scale_color_viridis(limits = c(1,2000),
                               breaks = c(1,1000,2000),
                               labels = c("1","1000","2000"))+
  # scale_color_gradient(na.value = 1)+
  # geom_smooth(method = "loess")+
  geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed", color = "black" )+
  # geom_smooth(method = "lm", color = "red", linetype = "dashed",linewidth = 1)+
  theme_Bede()+
  ylim(c(0.05,1))+
  xlim(c(0.05,1))+
  ylab("Standardised reflectance of the 120m flight")+
  xlab("Standardised reflectance of the 12m flight")+
  labs(color = "Pixel density")+
  theme(
    legend.position = "top",
        legend.key.width = unit(3,"line"))

@@

```{r ref_compare_building}
#| cache: true
#| echo: false
#| eval: false
#| warning: false

RAW <- ggplot()+
  geom_spatraster_rgb(data = rast("./Figures/High_res/Compare_reflectance.png"))+
  coord_equal()+
  geom_label(aes(x = 300,y= 1200, label = "A") ,size = 10)+
    theme_void()+
  theme(axis.text = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.ticks.length = unit(0, "pt"),
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "pt"))

STD <- ggplot()+
  geom_spatraster_rgb(data = rast("./Figures/High_res/Compare_reflectance_std.png"))+
  coord_equal()+
  geom_label(aes(x = 300,y= 1200, label = "B") ,size = 10)+
    theme_void()+
  theme(axis.text = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.ticks.length = unit(0, "pt"),
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "pt"))

@@

There was a slight underestimation for raw reflectance values in the high-altitude flight, particularly for higher reflectance values (@fig-CompareRef A). Since both flights were conducted over vegetated areas, the highest reflectance values correspond to the infrared part of the spectrum. This difference was not present when the reflectance has been standardized (@eq-std ; @fig-CompareRef B).

@@

Each drone flight was used to produce a prediction map, as well as a probability map that indicates the model derived probability of the selected class for every pixel. The low-altitude flight conducted in Gafanha, Portugal, represented the site with the highest complexity (@fig-GafLow). Among the five vegetation classes on which the model was trained, four were present on this site, with green and red macroalgae mixed with a seagrass meadow. There was also benthic diatoms forming biofilms on bare sediment surface. Although the seagrass was solely composed of a single specie, *Zostera noltei*, various colors of this specie could be observed from dark green (corresponding to healthy leaves) to brown (when leaves are senescent or have an altered pigment composition). Regardless of the variation of color, the class Magnoliopsida (seagrass) was accurately predicted by the model (F1 score of 0.96 at that site).

@@

The high-altitude flight over Gafanha covered a total area of ~1 km² (@fig-GafHigh). A channel contouring a small island was masked in the prediction map. Most of vegetation area was classified as seagrass by the model, including patches with discolored leaves. Only a few pixels were classified as green macroalgae (F1 score of 0.55). Patches of red macroalgae were correctly classified (F1 score of 0.85). In the northern part of the site and near the land eadges, patches of the schorre angiosperm *Sporobolus maritimus* (syn. *Spartina maritima)* were misclassified, either as seagrass or as brown algae (F1 score of 0.77 and 0.71, respectively).

@@

```{r Figure 5 Gaf High plot}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: RGB orthomosaic (Left) and Prediction (Right) of the high altitude flight of Gafanha, Portugal. The total extent of this flight was about 1 km² with a resolution of 80 mm per pixel. Background colors indicate intertidal area (Light Green), land area (Light Grey) and water (Light Blue). The yellow outline shows the extent of the low altitude flight of Gafanha presented in @fig-GafLow. The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel size.
#| label: fig-GafHigh
#| fig-width: 10
#| fig-height: 8
#| out-width: "100%"

>>>>

plot_RAW <- RAW %>%  
  ggplot(aes(x=Low, y=High))+
  # geom_point(aes(color = as.factor(pred)))
  ggpointdensity::geom_pointdensity(adjust = .03)+
  viridis::scale_color_viridis(limits = c(0,6000),
                               breaks = c(0,2000,4000,6000),
                               labels = c("1","2000","4000","6000"))+
  # geom_smooth(method = "loess")+
  geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed", color = "black" )+
  # geom_smooth(method = "lm", color = "red", linetype = "dashed",linewidth = 1)+
  theme_Bede()+
  ylim(c(0.05,0.6))+
  xlim(c(0.05,0.6))+
  ylab("Reflectance of the 120m flight")+
  xlab("Reflectance of the 12m flight")+
  labs(color = "Pixel density")+
  theme(
    axis.text.x = element_text(size = 15),
    plot.background = element_rect(fill = 'white', color = NA), 
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 17),
    axis.title.y = element_text(size = 17),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 17, vjust = 1),
    legend.position = "top",
        legend.key.width = unit(3,"line"))

plot_STD <- STD %>%  
    ggplot(aes(x=Low, y=High))+
  # geom_point(aes(color = as.factor(pred)))
  ggpointdensity::geom_pointdensity(adjust = .03)+
  viridis::scale_color_viridis(limits = c(1,2000),
                               breaks = c(1,1000,2000),
                               labels = c("1","1000","2000"))+
  scale_y_continuous(position = "right",
                     limits = c(0.05,1))+
  # scale_color_gradient(na.value = 1)+
  # geom_smooth(method = "loess")+
  geom_abline(intercept = 0, slope = 1, linewidth = 1, linetype = "dashed", color = "black" )+
  # geom_smooth(method = "lm", color = "red", linetype = "dashed",linewidth = 1)+
  theme_Bede()+
  xlim(c(0.05,1))+
  ylab("Standardised reflectance of the 120m flight")+
  xlab("Standardised reflectance of the 12m flight")+
  labs(color = "Pixel density")+
  theme(
    axis.text.x = element_text(size = 15),
    plot.background = element_rect(fill = 'white', color = NA), 
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 17),
    axis.title.y = element_text(size = 17),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 17, vjust = 1),
    legend.position = "top",
        legend.key.width = unit(3,"line"))

@@

```{r ref_compare_building}
#| cache: true
#| echo: false
#| eval: false
#| warning: false
img_raw <- rast("./Figures/High_res/Compare_reflectance.png")

RAW <- ggplot()+
  geom_spatraster_rgb(data = img_raw)+
  coord_equal(expand = F)+
  geom_label(aes(x = 300,y= 1200, label = "A") ,size = 10)+
    theme_void()+
  theme(axis.text = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.ticks.length = unit(0, "pt"),
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "pt"))

STD <- ggplot()+
  geom_spatraster_rgb(data = rast("./Figures/High_res/Compare_reflectance_std.png"))+
  coord_equal(expand = F)+
  geom_label(aes(x = 300,y= 1200, label = "B") ,size = 10)+
    theme_void()+
  theme(axis.text = element_blank(), 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.ticks.length = unit(0, "pt"),
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "pt"))

@@

There was a slight underestimation of raw reflectance values in the high-altitude flight, particularly for higher reflectance values (@fig-CompareRef A). Since both flights were conducted over vegetated areas, the highest reflectance values correspond to the infrared part of the spectrum. This difference was not present when the reflectance has been standardized (@eq-std ; @fig-CompareRef B).

@@

Each drone flight was used to produce a prediction map, as well as a probability map that indicates the model-derived probability of the selected class for every pixel. The low-altitude flight conducted in Gafanha, Portugal, represented the site with the highest complexity (@fig-GafLow). Among the five vegetation classes on which the model was trained, four were present on this site, with green and red macroalgae mixed with a seagrass meadow. There were also benthic diatoms biofilms on sediment surface. Although the seagrass was solely composed of a single species, *Zostera noltei*, two colors of this species could be observed: dark green (corresponding to healthy leaves) and brown (when leaves are senescent or have an altered pigment composition). Regardless of the variation of color, the class Magnoliopsida (seagrass) was accurately predicted by the model (F1 score of 0.96 at that site).

@@

The high-altitude flight over Gafanha covered a total area of ~1 km² (@fig-GafHigh). A channel contouring a small island was masked in the prediction map. Most of the vegetation area was classified as seagrass by the model, including patches with brown leaves. Only a few pixels were classified as green macroalgae (F1 score of 0.55). Patches of red macroalgae were correctly classified (F1 score of 0.85). In the northern part of the site and near the land edges, patches of the schorre angiosperm *Sporobolus maritimus* (syn. *Spartina maritima)* were misclassified, either as seagrass or as brown algae (F1 score of 0.77 and 0.71, respectively).

@@

```{r Figure 5 Gaf High plot}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: RGB orthomosaic (Left) and Prediction (Right) of the high-altitude flight of Gafanha, Portugal. The total extent of this flight was about 1 km² with a resolution of 80 mm per pixel. Background colors indicate the intertidal area (Light Green), land area (Light Grey) and water (Light Blue). The red outline shows the extent of the low altitude flight of Gafanha presented in @fig-GafLow. The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel size.
#| label: fig-GafHigh
#| fig-width: 10
#| fig-height: 8
#| out-width: "100%"
